---
title: "Lab 4 Analysis"
author: "Stephen Cook"
date: "Oct 22 2018"
output:
  pdf_document: default
---

## Import packages


```{r, message=FALSE}
require(tidyverse)
```


## Import data


```{r}
invert <- read.csv(file = "biodiversity_data_distributed.csv", header = T, sep = ",")
```

## Calculations

Which stream is more diverse (using richness-based measurements)? Neils or Harris Creek? 

```{r}
alpha_div <- invert %>%
  select(SITE, REP, CODE) %>%
  group_by(SITE, REP) %>%
  summarise(ALPHA_R = n_distinct(CODE))

gamma_div <- invert %>%
  select(SITE, CODE) %>%
  group_by(SITE) %>%
  summarise(GAMMA_R = n_distinct(CODE))
```

```{r}
richness_diversity <- alpha_div %>%
  summarise(ALPHA_R = mean(ALPHA_R)) %>%
  full_join(gamma_div, by = "SITE") %>%
  mutate(BETA_R = ALPHA_R/GAMMA_R)
```

%EPT is a common and simple biotic integrity score (remember these taxa are commonly sensitive to anthropogenic stressors).


```{r}
ept_taxa <- invert %>%
  select(SITE, REP, ORDER, COUNT) %>% # select the columns you need
  group_by(SITE, REP, ORDER) %>% # put appropriate grouping variables here
  summarise(COUNT = sum(COUNT)) %>% # sum across the order (for orders with more than one genus)
  filter(ORDER == "Ephemeroptera" | ORDER == "Plecoptera" | ORDER == "Odonata") %>%
  group_by(SITE, REP) %>%
  summarise(EPT_COUNT = sum(COUNT))
```

```{r}
all_taxa <- invert %>%
  select(SITE, REP, ORDER, COUNT) %>% # select the columns you need
  group_by(SITE, REP) %>%
  summarise(ALL_COUNT = sum(COUNT))
```

```{r}
combined <- ept_taxa %>%
  full_join(all_taxa, by = c("SITE", "REP")) %>%
  mutate(PERCENT_EPT = (EPT_COUNT/ALL_COUNT)*100)
```

# Statistical tests

Is alpha-diversity significantly differ between the sites?

```{r}
neils <- alpha_div %>%
  filter(SITE == "neils") 

harris <- alpha_div %>%
  filter(SITE == "harris") 
```


```{r}
diversity_ttest <- t.test(neils$ALPHA_R, harris$ALPHA_R)

diversity_ttest
```
Does the % of EPT in the assemblage differ between the 2 sites?

```{r}
neils_ept <- combined %>%
  filter(SITE == "neils") 

harris_ept <- combined %>%
  filter(SITE == "harris") 
```


```{r}
ept_ttest <- t.test(neils_ept$PERCENT_EPT, harris_ept$PERCENT_EPT)

ept_ttest
```





