---
title: "Survivorship Plot"
author: "SCC"
date: "May 30, 2018"
output: html_document
---

```{r}
require(tidyverse)
require(cowplot)
```


```{r}
plot1 <- ggplot(mtcars, aes(wt, mpg)) +
  geom_blank() +
  theme_classic(base_size = 12) +
  scale_y_log10(limits = c(1,1000), breaks = c(1,10,100,1000),
                expand = c(0,0)) +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank()) +
  ylab("Number of survivors") +
  xlab("Age (in relative units)") +
  annotation_logticks(sides = "l")

plot1
```

```{r}
ggsave(plot1, device = "pdf", filename = "base_plot.pdf", height = 3, width = 5)
```


- Water backup (home)
- transportation expense (car)
- roadside coverage (car)
- actual cash value vs bluebook value (replacement costs)
- personal property depreciated like roof? replacement cost or assessed value?
- ask about safe driver discounts (conditions)
- ask about all discounts and if they expire after 1 year
- responsible payment discount??
- need roof depreciation schedule

