---
title: 'Biology 3103 : Ecology Laboratory'
author: "Stephen Cook & Dr. Robert Doyle"
date: "Spring 2019"
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    latex_engine: xelatex
    template: svm-latex-syllabus.tex
  word_document: default
  html_document:
    df_print: paged
email: stephen_cook@baylor.edu, robert_doyle@baylor.edu
fontfamily: mathpazo
fontsize: 11pt
geometry: margin=1in
header-includes: \linespread{1.05}
office: C453.R (Cook), C419 (Doyle)
officehours: By appointment, but generally around
classroom: A222
classhours: 2:00-6:15
---

# Course Objectives

The goal of this lab is to develop the student as a scientist by exposure to basic research methodology in the field of ecology. You will 

1. develop understanding of ecology as a scientific discipline, and

2. become familiar with some common methods used by ecologists (including field, laboratory,
and data handling practices).

# General information

Each lab will consist of the same general elements. You will (1) form hypotheses, (2) collect some data, (3) analyze that data, and (4) write-up the results. Ecology is synonymous with field research - barring dangerous conditions (e.g., lightning) we
will be outdoors if the planned lab is outdoors. Rain or shine! Wearing clothes not suited to field
conditions will not be considered a valid excuse to not participate in the lab, so dress appropriately.

In the lab, university policy is that you must wear closed-toe shoes and long pants. Pay attention
to the weather and dress appropriately for outdoor conditions! For the wet labs I would wear old
jeans or quick-dry pants/shorts. Shorts could be fine- but we may be wading through vegetation
(scratchy). Old tennis shoes are highly recommended. You will not be allowed to wade barefoot
(i.e., you will not participate in that lab, and lose participation points). More specific info will be
provided for each outdoor lab.

If you ever have questions or concerns, please do not hesitate to email either Stephen (preferably)
or Dr. Doyle. We generally respond quickly to emails, with the exception of the evening before lab reports are due (Stephen will politely ignore report questions emailed after 5:00pm the day before the assignment is due). Please get in touch if you need help outside of lab - we are around and available to meet.

## Laboratory/field activities

It is important that students be aware that participating in laboratory activities carries the risk, however small, of injury.  Should you become injured during activities in this class, be aware of the following:

- Notify your instructor immediately if you are injured or exposed to a chemical.  Ask your instructor to provide the chemical safety data sheet (SDS) to take with you.
- The Student Health Center can treat minor physical injuries such as cuts, burns, etc. The Health Center is open during campus business hours.  After-hours medical care should be sought at an off-campus emergency care facility.
- The Health Center does not provide emergency care for eye injuries or chemical exposures.  In such cases, treatment should be sought at an emergency care center.
- The nearest 24/7 emergency care center is Premier located at 221 N. Jack Kultgen Expy (254) 537-9452.  Other nearby ERs include BSWMC Hillcrest and Express ER.
- Depending on the type/extent of injury, an ambulance may be requested.
- Medical expenses incurred as a result of a laboratory injury, including ambulance transport, are the financial responsibility of the student.

# Grading

## Point breakdown

- **72 points** - 6 lab reports at 12 points each

- **18 points** - 6 R analyses at 3 points each

- **10 points** - 2 field trips at 5 points each

## Grading scale

Minuses (e.g. B-) will not be given.

```{r table2, fig.align = "left", echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
tabl <- # simple table creation here
"| Points        | Letter grade  |
|---------------|:-------------:|
| 90.0-100      | A |
| 87.0-89.9     | B+      |
| 80.0-86.9 | B |
| 77.0-79.9 | C+ |
| 70.0-76.9 | C |
| 60-69.9 | D |
| <59.9 | F |"

cat(tabl) # output the table in a format good for HTML/PDF/docx conversion
```

## Lab reports

See below.

## Analyses in R

R is a user-friendly programming language that scientists in a wide variety of fields use for data analysis, graphing, and report generation. If computer programming is completely foreign to you, don't worry! You will be given everything you need for this lab. 

We will be using RstudioCloud, which enables you to log-in online to conduct your analysis and create figures for your reports. This document will be due the same time as each Lab Report. Look to a separate handout for information.

## Field trips (2x)

Field trips will involve attendance as well as active participation, which may include, for example, taking notes or completing a worksheet.


# Course Policies

## Absences

As this is a lab (and thus an experienced-based) course, it is imperative that you make it to class. Keep in mind that
it is very important to be on time, as we will frequently be leaving the lab to collect data in the field. Tardiness is not
a valid excuse for missing class. Attendance policy is detailed in the Baylor University student handbook. You will
be automatically dropped from the class if you have 4 unexcused absences. Missing a lab is forfeiture of the points
available for that day. For labs that take up two days, missing one of the labs will result in a 50% deduction on the
lab report. For an absence to be excused, the instructor must receive documentation (e.g., doctor's or chaplain's
note). Students with excused absences may be given alternate assignments.

## Canvas

We will use Canvas to post data, report guidelines, readings, and communications about the lab. You will also use
Canvas to submit lab reports. Each report will automatically be submitted to *Turnitin*, and will be graded and scored on Canvas.
**Please review any comments/suggestions in order to improve your lab reports.** We suggest you check Canvas each
week before coming to class for any updated announcements.


## Cell phones

Please do not use your cell phone in lab unless instructed to do so. Used responsibly, they can be an excellent tool to
record data or steps in the experiment (pictures or notes). If you chose to bring your cell phone to the field, we are
not responsible for them!

## Academic integrity

Plagiarism or any form of cheating involves a breach of student-teacher trust. This means that any work submitted under your name is expected to be your own, neither composed by anyone else as a whole or in part, nor handed over to another person for complete or partial revision. Be sure to document all ideas that are not your own. Instances of plagiarism or any other act of academic dishonesty will be reported to the Honor Council and may result in failure of the course. Not understanding plagiarism is not an excuse. You may use online resources to study for this course, but you must do so in ways that are consistent with all aspects of the Baylor University Honor Code (see, specifically, Section III.C.12 and Section III.C.16). As a Baylor student, I expect you to be intimately familiar with all aspects of the Honor Code, which can be found at this link: http://www.baylor.edu/honorcode/.

Translation... If you use outside resources,
please cite them appropriately, and do not copy directly from sources or the lab handouts. Though we encourage you to work in partners/groups, you must analyze the data and write reports independently. **Do not copy from current/previous students' lab reports.** Please be advised that each report will be automatically submitted to *Turnitin* through Canvas and assigned a similarity score.


## Disabilities policy

Any student who needs academic accommodations related to a documented disability should inform me immediately at the beginning of the semester. You are required to obtain appropriate documentation and information regarding accommodations from the Office of Access and Learning Accommodation (OALA). Contact Information: (254) 710-3605 - Paul L. Foster Success Center, 1st floor on the East Wing of Sid Richardson.

## Title IX statement

Baylor University does not discriminate on the basis of sex or gender in any of its education or employment programs and activities, and it does not tolerate discrimination or harassment on the basis of sex or gender. This policy prohibits sexual and gender-based harassment, sexual assault, sexual exploitation, stalking, intimate partner violence, and retaliation (collectively referred to as prohibited conduct).  For more information on how to report, or to learn more about our policy and process, please visit www.baylor.edu/titleix. You may also contact the Title IX office directly by phone, (254) 710-8454, or email, TitleIX_Coordinator@baylor.edu.

The Title IX office understands the sensitive nature of these situations and can provide information about available on- and off-campus resources, such as counseling and psychological services, medical treatment, academic support, university housing, and other forms of assistance that may be available. Staff members at the office can also explain your rights and procedural options if you contact the Title IX Office. You will not be required to share your experience. If you or someone you know feels unsafe or may be in imminent danger, please call the Baylor Police Department (254-710-2222) or Waco Police Department (9-1-1) immediately.

\newpage

# Monday (section 01) tentative schedule

```{r table_2, echo=FALSE, message=FALSE, warnings=FALSE, results='asis', fig.align='left'}
tabl_2 <- # simple table creation here
"| Date        | Lab  | What is due? |
|---------------|:-------------| ------:|
| Jan 14     | Introduction, LAB 1 data collection | Nothing |
| Jan 21     | MLK Day| Nothing |
| Jan 28     | LAB 1: Cemetery demography / golden eagle lab| R primers & reading |
| Feb 4     | LAB 2: Population ecology (algal bioassay) | LAB 1 |
| Feb 11     | LAB 2: Analysis / discussion | Nothing |
| Feb 18     | Field trip no. 1 | LAB 2 |
| Feb 25     | LAB 3: Photosynthesis | Nothing |
| March 4    | LAB 3: Analysis / discussion | Nothing |
| March 11     | SPRING BREAK | Nothing |
| March 18     | LAB 4: Vegetation transects | LAB 3 |
| March 25     | LAB 4: Analysis / discussion | Nothing |
| April 1     | LAB 5: Biodiversity (field)| LAB 4 |
| April 8     | LAB 5: Biodiversity (lab, analysis/discussion) | Nothing |
| April 15     | Field trip no. 2 - Cameron Park | LAB 5 |
| April 22     | EASTER BREAK | Nothing |
| April 29     | LAB 6: Climate modeling | Nothing |"

cat(tabl_2) # output the table in a format good for HTML/PDF/docx conversion

```

\begin{flushright}

**Lab 6 will be due on Wednesday, May 8 at 5:00pm.**

\end{flushright}

# Tuesday (section 02) tentative schedule

```{r table_3, echo=FALSE, message=FALSE, warnings=FALSE, results='asis', fig.align='left', fig.cap="This is a test."}
tabl_3 <- # simple table creation here
"| Date        | Lab  | What is due? |
|---------------|:-------------| ------:|
| Jan 15     | Introduction, LAB 1 data collection | Nothing |
| Jan 22     | MLK DAY | Nothing |
| Jan 29     | LAB 1: Cemetery demography / golden eagle lab| R primers & reading|
| Feb 5     | LAB 2: Population ecology (algal bioassay) | LAB 1 |
| Feb 12     | LAB 2: Analysis / discussion | Nothing |
| Feb 19     | Field trip no. 1 | LAB 2 |
| Feb 26     | LAB 3: Photosynthesis | Nothing |
| March 5    | LAB 3: Analysis / discussion | Nothing |
| March 12     | SPRING BREAK | Nothing |
| March 19     | LAB 4: Vegetation transects | LAB 3 |
| March 26     | LAB 4: Analysis / discussion | Nothing |
| April 2     | LAB 5: Biodiversity (field)| LAB 4 |
| April 9     | DIADELOSO | Nothing |
| April 16     | LAB 5: Biodiversity (lab, analysis/discussion) | Nothing |
| April 23     | Field trip no. 2 - Cameron Park | LAB 5 |
| April 30     | LAB 6: Climate modeling | Nothing |"

cat(tabl_3) # output the table in a format good for HTML/PDF/docx conversion
```

\begin{flushright}

**Lab 6 will be due on Thursday, May 9 at 5:00pm.**

\end{flushright}

\newpage

# Lab report guidelines

Lab reports are due via electronic submission (Canvas) by the beginning of lab time on the due dates mentioned on the course schedule. It is the student's responsibility to turn in lab reports. Labs may be accepted up to one week late and will receive a 30% reduction in score (maximum 70%), but will not be eligible for a rewrite. No labs will be accepted more than one week late unless special circumstances are approved by TA or Professor. 

If you receive a failing score (< 70%) on a report, and you turned it in on time, you have the opportunity to rewrite and resubmit the report for a maximum of 70%. This rewrite will be due one week from when the report is handed back. If you choose to take advantage of this, please carefully consider the comments/corrections on your graded report, and ask questions if necessary!

## Style and format for written report

Mirror your reports off the peer-reviewed studies we read in class. Points may be taken off the lab report for poor grammar, syntax, spelling, etc. Additionally, please be concise in your writing - a hallmark of good scientific writing is finding a way to clearly convey an idea in as few words as possible.

## Introduction

Briefly give introductory material explaining the importance of the lab material (e.g., how it relates to ecology in general and why we care about the subject matter). Pay attention during the lab set up- the instructors will state most of what you need to know for this! At the end of the introduction, you must also include a specific null research hypothesis. You may also make a prediction about the Results based on information you've given in the Introduction - that is, you may state an alternate hypothesis.  This alternate hypothesis should say whether you think there will be an effect, and in what direction that effect will occur.

1.	Establish the question you are interested in with a general objectives statement:
*"The objective of this study was to determine the effectiveness of the Lab Report Guidelines in aiding student success."*

2. Turn that question into a hypothesis that can be nullified with some degree of certainty:
*"We tested the null hypothesis that students who read the Lab Report Guidelines, and those that do not, receive similar scores on their Lab Reports."*

3.	Determine a plausible alternative hypothesis that your data could support if the null hypothesis is rejected:
*"Our alternative hypothesis was that students who read the Guidelines would perform better than students who did not."*

## Methods

Begin with a general, overarching statement about what you did, then clearly summarize what was done in lab to address the hypotheses.  The KEY points of all experimental design, data collection, and data analysis should be included. Make sure to include details about any calculations and/or statistics you used, and how you used them.

## Results

Summarize the data collected and the results used to address the hypothesis. The results in the text (in paragraph form) should be graphically represented - you will use the R to generate your figures, **with appropriate labelling of axes and legends, if applicable.** Clearly state the results of the statistical test employed.

- Graphs and tables MUST have a detailed caption explaining the contents!
- Figures must have labeled axes that include the units (if appropriate) 

## Discussion

Explain if the null hypothesis stated in the introduction was rejected or supported (you should remind the reader what hypothesis you were testing). Describe what the Results mean in the context of your Introduction. While the Results section details *what* you found, the Discussion section should be dedicated to rationalizing *why* you think your data displayed these results.  

\newpage

```{r echo=FALSE, fig.cap="Grading rubric for Ecology Lab of Fall 2018. Each section will be scored by the criteria provided under 'Aspect being evaluated', and points will total to 12 points."}
knitr::include_graphics('Grading_rubric.pdf')
```